### 2020-09-11 ccsn.tex
Done with the introduction

%\documentclass[aps, twocolumn, superscriptaddress, showpacs, linenumbers, nofootinbib]{revtex4-1}
\documentclass[aps, twocolumn, superscriptaddress, showpacs, nofootinbib, longbibliography]{revtex4-1}

% \documentclass[aps,prl,reprint,showpacs]{revtex4-1}

%\usepackage{lipsum}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{ae,aecompl} 
\usepackage{graphicx}
\graphicspath{{figures/}} % Directory in which figures are stored

\usepackage{amsmath}
\usepackage{color}
\usepackage{amssymb}
\usepackage{latexsym}
\usepackage{wasysym}
\usepackage{psfrag}
%\usepackage{comment}
\usepackage{ifthen}
\usepackage[citecolor=blue,colorlinks=true]{hyperref}
\usepackage{longtable}
\usepackage{float}
\usepackage[utf8]{inputenc}
\usepackage{lineno}
\usepackage{units}
\usepackage[title]{appendix}
\usepackage{aas_macros}
\usepackage{ulem}
\usepackage{subfigure}
\newcommand{\note}[1]{{\color{red}#1}}
\newcommand{\second}{\ensuremath{\mathrm{\,s}}}
%\newcommand{\mnras}{MNRAS}
\def\apjl{Astrophys. J. Lett.}
\def\prd{Phys.~Rev.~D}
\def\aap{A\&A}

\definecolor{maob}{rgb}{0.6,0.1,0.2}
\newcommand{\maob}[1]{\textcolor{maob}{#1}}
\newcommand{\pcd}[1]{\textcolor{blue}{#1}}

\newcommand{\Msol}{M_{\odot}}

% \usepackage{svn-multi}
%\restylefloat{table}

\begin{document}
\linenumbers

\title{Identification of protoneutron star g-modes in gravitational-wave data
\pcd{[PCD: alt. title]Inference of proto-neutron star properties from gravitational wave data in core-collapse supernovae.}}

\begin{abstract}

\end{abstract}

\maketitle

\input{introduction}\bigskip

\input{simulations}\bigskip

\input{methods}\bigskip

\input{results}\bigskip

\input{discussion}\bigskip

\bigskip\noindent\textit{Acknowledgments} ---

\begin{appendices}
\input{appendixA}
\end{appendices}
%\bibliographystyle{unsrt}
\bibliography{biblio}

\end{document}