### 2020-10-13 ccsn.tex
add NLC comments

%\documentclass[aps, twocolumn, superscriptaddress, showpacs, linenumbers, nofootinbib]{revtex4-1}
\documentclass[aps, twocolumn, superscriptaddress, showpacs, nofootinbib, longbibliography]{revtex4-1}

% \documentclass[aps,prl,reprint,showpacs]{revtex4-1}

%\usepackage{lipsum}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{ae,aecompl} 
\usepackage{graphicx}
\graphicspath{{figures/}} % Directory in which figures are stored

\usepackage{amsmath}
\usepackage{color}
\usepackage{amssymb}
\usepackage{latexsym}
\usepackage{wasysym}
\usepackage{psfrag}
%\usepackage{comment}
\usepackage{ifthen}
\usepackage[citecolor=blue,colorlinks=true]{hyperref}
\usepackage{longtable}
\usepackage{float}
\usepackage[utf8]{inputenc}
\usepackage{lineno}
\usepackage{units}
\usepackage[title]{appendix}
\usepackage{aas_macros}
\usepackage{ulem}
\usepackage{subfigure}
\usepackage{multirow}
%\newcommand{\note}[1]{{\color{red}#1}}
\newcommand{\second}{\ensuremath{\mathrm{\,s}}}
%\newcommand{\mnras}{MNRAS}
\def\apjl{Astrophys. J. Lett.}
\def\prd{Phys.~Rev.~D}
\def\aap{A\&A}

\definecolor{maob}{rgb}{0.6,0.1,0.2}
\newcommand{\maob}[1]{\textcolor{maob}{#1}}
\newcommand{\pcd}[1]{\textcolor{blue}{#1}}
\newcommand{\mab}[1]{\textcolor{red}{#1}}

\newcommand{\Msol}{M_{\odot}}

% \usepackage{svn-multi}
%\restylefloat{table}

\begin{document}
\linenumbers

\title{Inference of proto-neutron star properties from gravitational wave data in core-collapse supernovae.}

\author{M.~A.~Bizouard}
\affiliation{Artemis, Universit\'e C\^ote d'Azur, Observatoire C\^ote d'Azur, CNRS, CS 34229, F-06304 Nice Cedex 4, France}

\author{P.~Cerd\'a-Dur\'an}
\affiliation{Departamento de Astronom\'{\i }a y Astrof\'{\i }sica, Universitat de Val\`encia, E-46100 Burjassot, Val\`encia, Spain}

\author{N.~Christensen}
\affiliation{Carleton College, Northfield, MN 55057, USA}
\affiliation{Artemis, Universit\'e C\^ote d'Azur, Observatoire C\^ote d'Azur, CNRS, CS 34229, F-06304 Nice Cedex 4, France}

\author{J.~A.~Font}
\affiliation{Departamento de Astronom\'{\i }a y Astrof\'{\i }sica, Universitat de Val\`encia, E-46100 Burjassot, Val\`encia, Spain}

\author{P.~Maturana Russel}
\affiliation{Department of Statistics, The University of Auckland, Auckland, New Zealand}
\affiliation{Department of Mathematical Sciences, Auckland University of Technology, Auckland, New Zealand}

\author{R.~Meyer}
\affiliation{Department of Statistics, The University of Auckland, Auckland, New Zealand}

\author{M.~Obergaulinger}
\affiliation{Departamento de Astronom\'{\i }a y Astrof\'{\i }sica, Universitat de Val\`encia, E-46100 Burjassot, Val\`encia, Spain}

\author{A.~Torres-Forn\'e}
\affiliation{Max Planck Institute for Gravitationalphysik (Albert Einstein Institute), D-14476 Potsdam-Golm, Germany}
\affiliation{Departamento de Astronom\'{\i }a y Astrof\'{\i }sica, Universitat de Val\`encia, E-46100 Burjassot, Val\`encia, Spain}

\mab{Alphabetical order ? Better to move forward the youngest}

\begin{abstract}
  Core collapse supernovae are very important phenomena in the Universe and their expected gravitational
  waves emission is a promising tool to study the onset of the supernova explosion mechanism. The complexity
  of the gravitational wave signal due to matter effects and the large degrees of freedom of the phenomena
  makes the source parameter inference problem very challenging. In this paper, we have considered the
  proto neutron star oscillation modes that carry out a large fraction of the gravitational wave signal and
  whose frequency time evolution is linked to the system properties through universal relations as
  demonstrated in ~\cite{Torres:2019b}. We have developed a simple algorithm to extract from the
  gravitational wave data the frequency evolution of the gravity mode. A set of 1D core collapse
  supernova simulations is used to build a model of evolution of the proto neutron star properties
  with the gravity mode frequency. The model is then used to infer the time evolution of a combinaison of
  the radius and mass of the proto neutron star. We have then estimated the performance of the method by
  performing simulations with 2D core collapse supernova waveforms covering a progenitor mass range between
  11 and 40 solar mass and different equations of state. We have shown that for Advanced LIGO and Advanced
  Virgo detectors at design sensitivity it will be possible to infer proton neutron star properties for
  a galactic source, while third generation detectors Einstein Telescope and Cosmic Explorer will allow
  to test distances of several hundreds of kiloparsec.

\end{abstract}

\maketitle

\input{introduction}\bigskip

\input{simulations}\bigskip

\input{methods}\bigskip

\input{results}\bigskip

\input{conclusion}\bigskip

\bigskip\noindent\textit{Acknowledgments} ---

\begin{appendices}
\input{appendixA}
\end{appendices}
%\bibliographystyle{unsrt}
\bibliography{biblio}

\end{document}